---
title: "Homework #5: Density Estimation" 
author: "**Hyunsuk Ko**"
date: "Due: Wed Oct 12 | 11:45am"
output: R6030::homework
---

**DS 6030 | Fall 2022 | University of Virginia **

*******************************************
```{r config, echo=FALSE}
source(system.file("config/hw_config.R", package="R6030")) # knitr settings
options(dplyr.summarise.inform = FALSE)  # ignore dplyr message about grouping
```

# Required R packages and Directories

::: {.solution}
```{r packages, message=FALSE, warning=FALSE}
data.dir = 'https://mdporter.github.io/DS6030/data/' # data directory
library(R6030)     # functions for DS 6030
library(ks)        # functions for KDE
library(tidyverse) # functions for data manipulation   
```
:::



# Problem 1 Geographic Profiling

```{r, echo=FALSE, eval=FALSE}
set.seed(2019)
n = 283
sd = 2.1
x = sqrt(rnorm(n, sd=sd)^2 + rnorm(n, sd=sd)^2)

geo = read.csv("../data/geo_profile.csv", header = FALSE) %>% mutate(dist = V1)
geo = geo %>% dplyr::select(dist)
```

Geographic profiling, a method developed in criminology, can be used to estimate the [home location (roost) of animals](https://www.sciencedirect.com/science/article/pii/S0022519305004157) based on a collection of sightings. The approach requires an estimate of the distribution the animal will travel from their roost to forage for food. 

A sample of $283$ distances that pipistrelle bats traveled (in meters) from their roost can be found at: 

- **Bat Data**: <`r file.path(data.dir, 'geo_profile.csv')`>

One probability model for the distance these bats will travel is:
\begin{align*}
f(x; \theta) = \frac{x}{\theta} \exp \left( - \frac{x^2}{2 \theta} \right)
\end{align*}
where the parameter $\theta > 0$ controls how far they are willing to travel. 


## a. Derive the MLE for $\theta$ (i.e., show the math). 

::: {.solution}
$L = \frac{x}{\theta} \exp \left( - \frac{x^2}{2 \theta} \right)$

$logL$ = $\sum_{i=1}^{n} log (\frac{x}{\theta} * \exp \left( - \frac{x^2}{2 \theta} \right)) = \sum_{i=1}^{n} logx - log\theta + log(\exp \left( - \frac{x^2}{2 \theta} \right))$
:::


## b. What is the MLE of $\theta$ for the bat data? (Use results from a, or use computational methods.) 

::: {.solution}
$\frac{\partial logL}{\partial \theta} = \sum_{i=1}^{n} -\frac{1}{\theta} + \dfrac{1}{\exp \left( - \frac{x^2}{2 \theta} \right)} * \exp \left( - \frac{x^2}{2 \theta} \right) * \dfrac{x^2}{2\theta^2} = \sum_{i=1}^{n}  -\frac{1}{\theta} + \dfrac{x^2}{2\theta^2} = -\dfrac{n}{\theta} + \dfrac{n}{2\theta^2} * \sum_{i=1}^{n} x^2_i = \dfrac{-2n\theta + n * \sum_{i=1}^{n} x^2_i}{2\theta^2} = 0$


$\therefore \theta = \dfrac{\sum_{i=1}^{n} x^2_i}{2n}$
:::

## c. Using the MLE value of $\theta$ from part b, compute the estimated density at a set of evaluation points between 0 and 8 meters. Plot the estimated density.

::: {.solution}
```{r}
mle_theta = sum(geo$dist ^ 2)/ (2*n)

function_x = function(x) {
  result = (x / mle_theta) * exp(-(x^2) / (2 * mle_theta))
  return (result)
}

ggplot() +
  xlim(0,8) +
  geom_function(fun = function_x) +
  labs(title = "Density Plot of MLE")
```

:::

## d. Estimate the density using KDE. Report the bandwidth you chose and produce a plot of the estimated density. 

::: {.solution}
```{r}
bw = 5
bks = seq(0,8)
hh = hist(geo$dist, breaks = bks)

hh

f = kde(geo$dist, h = bw/3)
plot(f)

plot(hh,freq=FALSE,ylim=c(0,max(c(hh$density,f$estimate))), las=1,main='',border='white',col='grey75') 
rug(jitter(geo$dist)) 
lines(f$eval.points,f$estimate,col=2,lwd=1.25)
```

:::




## e. Which model do you prefer, the parametric or KDE? 

::: {.solution}
As KDE gives similar distribution as the true distribution, I prefer KDE.
:::




# Problem 2: Interstate Crash Density

Interstate 64 (I-64) is a major east-west road that passes just south of Charlottesville. Where and when are the most dangerous places/times to be on I-64? The crash data (link below) gives the mile marker and fractional time-of-week for crashes that occurred on I-64 between mile marker 87 and 136 in 2016. The time-of-week data takes a numeric value of *\<dow\>.\<hour/24\>*, where the dow starts at 0 for Sunday (6 for Sat) and the decimal gives the time of day information. Thus `time=0.0417` corresponds to Sun at 1am and `time=6.5` corresponds to Sat at noon. 

- **Crash Data**: <`r file.path(data.dir, 'crashes16.csv')`>

```{r}
#readr::write_csv(tibble(x), "../data/crashes16.csv", col_names=FALSE)

crash = read.csv("../data/crashes16.csv")
crash
```


## a. Extract the crashes and make a scatter plot with mile marker on x-axis and time on y-axis. 

::: {.solution}
```{r}
ggplot(crash, aes(x = mile, y = time)) +
  geom_point()
```
:::


## b. Use KDE to estimate the *mile marker* density. 

- Report the bandwidth. 
- Plot the density estimate.

::: {.solution}
```{r}
bw = 5
bks = seq(0, 136)
hh = hist(crash$mile, breaks = bks)

f = kde(crash$mile, h = bw/3)

plot(hh,freq=FALSE,ylim=c(0,max(c(hh$density,f$estimate))), las=1,main='',border='white',col='grey75') 
rug(jitter(crash$mile)) 
lines(f$eval.points,f$estimate,col=2,lwd=1.25)
```

:::


## c. Use KDE to estimate the temporal *time-of-week* density. 

- Report the bandwidth. 
- Plot the density estimate. 

::: {.solution}
```{r}
bw = 10
bks = seq(0, 7)
hh = hist(crash$time, breaks = bks)

f = kde(crash$time, h = bw/3)

plot(hh,freq=FALSE,ylim=c(0,max(c(hh$density,f$estimate))), las=1,main='',border='white',col='grey75') 
rug(jitter(crash$time)) 
lines(f$eval.points,f$estimate,col=2,lwd=1.25)
```

:::


## d. Use KDE to estimate the bivariate mile-time density. 

- Report the bandwidth parameters.
- Plot the bivariate density estimate. 

::: {.solution}
```{r}
H1 = Hscv(crash)

#smoothed cross-validation bw estimator 
f1 = kde(crash,H=H1)

#useHformultivariatedata 
plot(f1, cont=c(10,50,95), las=1,xlim=c(85,140),ylim=c(-1,8))
points(crash,pch=19,cex=.5,col='grey60')
grid()
```

:::


## e. Based on the estimated density, approximate the most dangerous place and time to drive on this strech of road. Identify the mile marker and time-of-week pair. 

::: {.solution}
Mile between 100 ~ 115 at Friday night, and mile between 115 ~ 120 at Monday night seems to be the most dangerous place and time to drive.
:::
		